Analyzing the impact of social attributes on the success of commit integration

### 2017-02-19 Paper/MSRChallenge2017.tex

\usepackage[usenames,dvipsnames]{color}
\newcommand{\todo}[1]
  {{\scriptsize \textbf{\color{red} {#1}}}}

@@

\section{Introduction}
For managers of an open source project, it is important to ensure that the
project consists of high quality contributions.  The decision to include code in
a project has been described as a meritocracy, where only the code is a factor
in the decision~\cite{Scacchi07}.  However, to better understand successful 
project contributions, we must take into consideration and analyze 
different social attributes from the developers creating the commits,
which can be great predictors for the success of a contribution.

@@

\section{Research Design}
We present the approach taken to guide this study by
following the Goal-Question-Metric (GQM) methodology descibed by Basili et 
al~\cite{Basili84}.

\subsection{Goal}
The goal of this study is to have a deeper understanding of the relationship 
between the several different social properties that estimate the
developer's experience, and the likelihood of a developer's commit
to be considered successfully integrated in the project. We define successful
integration as passing the continuous integration build tests in the project.

\subsection{Question}

Previous studies have identified a relationship between technical and social 
factors and pull request acceptance. We are interested in understanding if
social factors, used as a proxy for experience, can be used to predict
a different success measure: commit integration success, therefore giving
developers and integrators another tool to be able to identify the quality
of a contribution performed by a particular developer.

@@

\subsection{Metric}
We created a decision tree based on two different sets of data. Merely 
technical data related to technical aspects of the complexity of the commit, and 
a superset of this before-mentioned data and social data gathered from the 
developer who pushed the commit. We evaluate the predictive power of these two 
models with 10 fold cross validation and we analyze which of them is able to 
predict with higher accuracy the success of unseen commits.

\section{Building Dataset}
\label{approach}
To investigate our research questions, we built a data set based on the 
attributes that consisted of both technical and social factors. For the technical factors, we collected
these from the data set provided by the MSR Challenge~\cite{msr17challenge}.
We extracted only the attributes in the data set that proxy the commit's
complexity and the build status as listed below:
\begin{itemize}
%\item tr\_original\_commit
%\item gh\_build\_started\_at 
\item git\_diff\_src\_churn 
\item gh\_diff\_files\_added 
\item gh\_diff\_files\_deleted 
\item gh\_diff\_files\_modified 
\item gh\_diff\_tests\_added 
\item gh\_diff\_tests\_deleted 
\item gh\_diff\_src\_files 
\item gh\_sloc 
\item tr\_status 
\end{itemize}

@@

\hline
\end{tabular}
\\
\center
  \caption{ Second and third columns: Comparison of performance when running 10 fold cross validation of two decision trees. The first column after the first double bar uses only technical attributes to predict the Travis Job status, the second column augments the data set with social attributes. Fourth and fifth columns: The same intuition as the previous two, with the difference that the data set was downsampled to provide the same number of successful and unsuccessful instances }
  \label{resultsTable}
\end{table*}

As a result of extracting just these particular technical attributes, the new data set is conformed by a  big number of repeated rows, which would slow down the process of collecting the social data for each of these rows, therefore, at this point we removed the repeated rows and continued to collect the social attributes from the developers who pushed the commits.

To collect the social data, we created a web scrapper to extract the social data
from developer's profiles.
We used the commit hash and the project name in the original data
set provided by the MSR Challenge~\cite{msr17challenge} to locate the web page of the commit using the syntax
below. 
\newline \newline
\textit{https://github.com/\textbf{NameOfProject}/commit/\textbf{CommitHash}}
\newline \newline

Then, we scraped the commit author's
\textit{username} from the page. If the scrapper was able to find username, the
scraper went to the user's profile information using the syntax below.
\newline \newline
\textit{https://github.com/\textbf{UserName}}
\newline \newline
The scrapper then extracted the following items from the user's profile.
\begin{itemize}
\item Repositories
\item Stars
\item Followers
\item Following
\item Contributions in the last year
\end{itemize}

We also added one item from the original data set to the social factors:
gh\_by\_core\_team\_member.
This factor indicates if the developer making a contribution is a core member 
of that project.

We used this data to create two data sets.  One with only the technical
factors and the build status, and another with both the technical and social
factors along with the build status.

The data set used in this study~\cite{msr17challenge} is very diverse 
and consists of commits that date from far back, and the social data we are 
gathering is recent data, gathered in January 2017. The researchers of this 
project acknowledge this discrepancy and in an effort to mitigate the threat of
using future result to predict previous events,
we removed the commits in the data set that were not performed within the last
three years of the data set's release (2014,2015,2016).  While this is not
perfect, we believe that these numbers are unlikely to have changed
significantly in the last three years.

After building the data set, we converted the various Travis build statuses to
successful (passed in the data set) or not successful (errored, failed, canceled
and started in the data set).  This still left a big discrepancy in the two
categories, with about a 3:1 successful to not successful ratio. To avoid having
this discrepancy alter our results, we created a separate data set where the
successful builds were randomly selected until the number of successful builds
equaled the number of not successful builds (downsampling process).  We used both the
downsampled and non-downsampled data sets in our results.

@@

\section{Results}

@@

In the original data set, 74.43\% of the builds were successful, the rest were not
successful.  This means that a classifier which guesses successful every time
would produce a classification rate of 74.43\% if ran on the full data set.

Table~\ref{resultsTable} shows the results of running the two data sets through the
default Weka J48 decision tree algorithm. The decision tree built from
only the technical indicators has a correct classification rate of 77.01\% which
means the technical attributes only increased the prediction accuracy by about 3\%.
Adding the social data increased the classifier's accuracy from 77.01\% to
81.10\%, about a 4\% increase.

@@

@@

For the build success investigation, we classified the build success rate for
core team members and non-core team members.  We found that core team members
triggered of 89.39\% of the builds in the data set and 75.58\% of those builds
passed.  Non-core team members triggered 10.61\% of the builds in the data set
and 64.70\% of the builds passed.

@@

\section{Discussion}
We have found that social attributes which proxy experience produce a 
noticeable increase in predicting build success.  While the increase was noticeable 
with the full data set, there was a much greater increase in prediction
accuracy after performing downsampling. These results show that social factors
as the ones studied in this paper are important predictors when
evaluating the predictive power of build commit success.

@@

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

\section{Threats to validity}
\noindent\textbf{Internal validity:}\\
The researchers in this study used the dataset provided by the 
 MSRChallenge~\cite{msr17challenge} to gain useful information regarding the 
 behavior of developers when pushing commits and how well these commits integrate 
 successfully with the project. The authors have used 
 Weka,\footnote{http://www.cs.waikato.ac.nz/ml/weka/} the well known tool for 
 usage of machine learning algorithms to analyze our datasets. We are aware that, 
 as every human written tool, Weka is prone to contain errors and/or bugs; 
 nonetheless, Weka is a broadly used, open source tool that is subject to high 
 quality standards and a vast user and developer base which makes it more likely 
 to be a high quality code and to have less bugs than lesser used tools.

@@

\noindent\textbf{External validity:} \\
 To tackle the concern that the results of this study might not generalize to 
 other instances outside of the studied cases, we have used a broad and rich 
 dataset~\cite{msr17challenge} in our analysis that comprises 1,359 projects. The 
 projects used in this dataset have also been filtered to be popular (having more 
 than 10 watches on Github), and to have a history of using the continuous 
 integration tool Travis CI for more than 50 builds. All these features help to 
 build the notion that since we are building our analysis on top of a vast 
 dataset comprehending a wide diversity of projects, the threat of this analysis 
 not generalizing to other projects is mitigated.

\noindent\textbf{Construct validity:}\\
The authors understand that 
 Travis job success is not the only way to analyze the success of a commit, and 
 that for different reasons the job may not be successful even if the code 
 commited does not have flaws (a flacky test case for example). Since continuous 
 integration is a best known practice in industry and the use of this methodology 
 spreads out into the open source community, we consider that cases of misuse of 
 CI such that it will lead to failing travis jobs when they should have actually 
 be successful are minimum, specially since having a solid build is one of the 
 main purposes of the methodology, and taking into account that the projects in 
 out dataset are projects with a history of using continuous integration and 
 having a certain level of popularity, this threat is minimized. We therefore 
 consider Travis job success to be a proxy 
 for success when integrating a commit into a larger project.

@@

>>>>

\usepackage[usenames,dvipsnames]{color}
\newcommand{\todo}[1]
  {{\scriptsize \textbf{\color{red} {#1}}}}
\newcommand{\clg}[1]
  {{\scriptsize \textbf{\color{blue} {CLG says: #1}}}}

@@

\section{Introduction}
\clg{detail nitpick: ``et al.'' has a paragraph after the al. (and not the et).
  This is something I write about in the blog post you all claim to have read.
  Please fix it throughout.}

For managers of an open source project, it is important to ensure that the
project consists of high quality contributions.  The decision to include code in
a project has been described as a meritocracy, where only the code is a factor
in the decision~\cite{Scacchi07}.  However, to better understand successful 
project contributions, we must take into consideration and analyze 
different social attributes from the developers creating the commits,
which can be great predictors for the success of a contribution.

@@

\section{Research Design}

We present the approach taken to guide this study by
following the Goal-Question-Metric (GQM) methodology descibed by Basili et 
al~\cite{Basili84}.

\paragraph{Goal}
The goal of this study is to identify and understand relationships 
between the several different social properties that estimate the
developer's experience, and the likelihood of a developer's commit
to be considered successfully integrated in the project. We define successful
integration as passing the continuous integration build tests in the project.
\clg{I might recommend a paragraph at the top of this section defining at a high
  level the key terms, like ``technical elements'' versus ``social factors.''
  You do concretize them in the next section, but conceptually here they're a
  little bit UBD.}

\paragraph{Question}
Previous studies have identified a relationship between technical and social 
factors and pull request acceptance. We are interested in understanding if
social factors, used as a proxy for experience, can be used to predict
a different success measure: commit integration success, therefore giving
developers and integrators another tool to be able to identify the quality
of a contribution performed by a particular developer.

@@

\paragraph{Metric}
\clg{I think I need a little bit more detail here, at multiple levels (like, why
  a decision tree? among other questions.}
We created a decision tree based on two different sets of data. Merely 
technical data related to technical aspects of the complexity of the commit, and 
a superset of this before-mentioned data and social data gathered from the 
developer who pushed the commit. We evaluate the predictive power of these two 
models with 10 fold cross validation and we analyze which of them is able to 
predict with higher accuracy the success of unseen commits.

\section{Building Dataset}
\label{approach}

\clg{Perhaps it makes sense to break this down into subsections?}

To investigate our research questions, we built a data set based on the 
attributes that consisted of both technical and social factors. For the technical factors, we collected
these from the data set provided by the MSR Challenge~\cite{msr17challenge}.
We extracted only the attributes in the data set that proxy the commit's
complexity and the build status as listed below:\clg{The listed terms don't mean
  anything on their own; they're clearly grabbed from code.  Please rename them
  to something descriptive (in English) and give a quick definition of each.
  Even better is if you give some insight as to why they might be meaningful in
  your theory (or not).}
\begin{itemize}
%\item tr\_original\_commit
%\item gh\_build\_started\_at 
\item git\_diff\_src\_churn 
\item gh\_diff\_files\_added 
\item gh\_diff\_files\_deleted 
\item gh\_diff\_files\_modified 
\item gh\_diff\_tests\_added 
\item gh\_diff\_tests\_deleted 
\item gh\_diff\_src\_files 
\item gh\_sloc 
\item tr\_status 
\end{itemize}

As a result of extracting just these particular technical attributes, the new
data set is conformed by a  big number of repeated rows, which would slow down
the process of collecting the social data for each of these rows, therefore, at
this point we removed the repeated rows and continued to collect the social
attributes from the developers who pushed the commits.\clg{I'm afraid this
  sentence/paragraph isn't parsing for me.  Please rephrase, clarify, expand,
  and add precision.  Also break into multiple sentences, because it's really
  really long, which doesn't help.  To start, I think the word ``conformed'' is
  not what you mean.}

To collect the social data, we created a web scraper to extract the social data
from developer's profiles.
We used the commit hash and the project name in the original data
set provided by the MSR Challenge~\cite{msr17challenge} to locate the web page
of the commit, leveraging \textit{github.com}'s URL standard for projects and
commit hashes.
%  using the syntax
% below. 
% \newline \newline
% \textit{https://github.com/\textbf{NameOfProject}/commit/\textbf{CommitHash}}
% \newline \newline

We scraped the commit author's
\textit{username} from a commit page. If the scrapper\clg{you mean ``scraper''
  not ``scrapper''; I fixed it once but you should fix it everywhere else.} was able to find username, the
scraper went to the user's profile information
(\textit{https://github.com/\textbf{UserName}}).

The scrapper then extracted the following items from the user's
profile.\clg{slightly better than the above for technical factors because you do
  define them in the text below but I think integrating into the itemized list
  would be better.  Also, what does each of these mean?}
\begin{itemize}
\item Repositories
\item Stars
\item Followers
\item Following
\item Contributions in the last year
\end{itemize}

We also added one item from the original\clg{...what?} data set to the social factors:
gh\_by\_core\_team\_member.
This factor indicates if the developer making a contribution is a core member 
of that project.

We used this data to create two data sets.  One with only the technical
factors and the build status, and another with both the technical and social
factors along with the build status.\clg{This is not a complete
  sentence.}\clg{also: why do you build two data sets?}

The data set used in this study~\cite{msr17challenge} is very diverse 
and consists of commits that date from far back. The social data we 
gather is recent, gathered in January 2017. In an effort to mitigate the threat of
using future result to predict previous events,
we removed the commits in the data set that were not performed within the last
three years of the data set's release (2014,2015,2016).  While this is not
perfect, we believe that these numbers are unlikely to have changed
significantly in the last three years.\clg{...did you spot check?}

After building the data set, we converted the\clg{which?} various Travis build statuses to
successful (passed in the\clg{which?} data set) or not successful (errored, failed, canceled
and started in the data set).  This still left a big discrepancy in the two
categories, with about a 3:1 successful to not successful ratio.\clg{I think I'm
  having ah ard time here because it's not clear which dataset we're referring
  to, but I don't understand the previous sentence, and thus don't understand
  the next sentence.} To avoid having
this discrepancy alter our results, we created a separate data set where the
successful builds were randomly selected until the number of successful builds
equaled the number of not successful builds (downsampling process).\clg{How?}  We used both the
downsampled and non-downsampled data sets in our results.\clg{Why?}

@@

\hline
\end{tabular}
\\
\center
  \caption{ Second and third columns: Comparison of performance when running 10
    fold cross validation\clg{We've discussed this phrasing surrounding 10-fold cross
      validation.  Please fix.  What is the metric here?  How do I interpret
      these numbers?} of two decision trees. The first column after
    the first double bar uses only technical attributes to predict the Travis
    Job status, the second column augments the data set with social
    attributes. Fourth and fifth columns: \clg{this sentence is not a
      sentence/is a bit awkward.} The same intuition as the previous
    two, with the difference that the data set was downsampled to provide the
    same number of successful and unsuccessful instances.\clg{...and what is the
    takeaway from this table?} \clg{Small thing: labels should be inside the
    caption otherwise the numbers sometimes get confused.  It's a stupid latex thing.}  \label{resultsTable}} 
\end{table*}

@@

\clg{I killed the separation between ``evaluation'' and ``results'' because I
  don't know why it existed in the first place.}
\clg{I also wouldn't mind an enumeration of research questions; the stuff in
  design (section 2) is high level and it's not obvious where you're going to go
  at this juncture.}

@@

In the original data set, 74.43\% of the builds were successful, the rest were not
successful.  This means that a classifier which guesses successful every time
would produce a classification rate of 74.43\% if ran on the full data
set. \clg{This sentence migth be better place above, when you first discuss subsampling.}

Table~\ref{resultsTable} shows the results of running the two data sets through the
default Weka J48 decision tree algorithm. The decision tree built from
only the technical indicators has a correct classification rate of 77.01\% which
means the technical attributes only increased the prediction accuracy by about 3\%.
Adding the social data increased the classifier's accuracy from 77.01\% to
81.10\%, about a 4\% increase.\clg{It's not clear from this text that you're
  talking about the unsampled data.}

@@

\clg{...upshot? Also: there are various numbers in the Table that aren't
  discussed or explained either in the caption or in these results, like, what's
  a Kappa statistic, and what do we learn from them?}

@@

For the build success investigation,\clg{Is this a different investigation than
  the previous subsection?}  we classified the build success rate for
core team members and non-core team members.  We found that core team members
triggered of 89.39\% of the builds in the data set and 75.58\% of those builds
passed.  Non-core team members triggered 10.61\% of the builds in the\clg{which?} data set
and 64.70\% of the builds passed.

\clg{upshot?}

@@

\clg{upshot?}

\subsection{Discussion}

Overall, we have found that social attributes which proxy experience produce a 
noticeable increase in predicting build success.  While the increase was noticeable 
with the full data set, there was a much greater increase in prediction
accuracy after performing downsampling. These results show that social factors
as the ones studied in this paper are important\clg{...maybe a strong word?} predictors when
evaluating the predictive power of build commit success.

@@

\clg{connection back to theory?}

\section{Threats to validity}

\paragraph{Internal validity:}
The researchers in this study used the dataset provided by the 
 MSRChallenge~\cite{msr17challenge} to gain useful information regarding the 
 behavior of developers when pushing commits and how well these commits integrate 
 successfully with the project. The authors have used 
 Weka,\footnote{http://www.cs.waikato.ac.nz/ml/weka/} the well known tool for 
 usage of machine learning algorithms to analyze our datasets. We are aware that, 
 as every human written tool, Weka is prone to contain errors and/or bugs; 
 nonetheless, Weka is a broadly used, open source tool that is subject to high 
 quality standards and a vast user and developer base which makes it more likely 
 to be a high quality code and to have less bugs than lesser used tools.

@@

\paragraph{External validity:} 
 To tackle the concern that the results of this study might not generalize to 
 other instances outside of the studied cases, we have used a broad and rich 
 dataset~\cite{msr17challenge} in our analysis that comprises 1,359 projects. The 
 projects used in this dataset have also been filtered to be popular (having more 
 than 10 watches on Github), and to have a history of using the continuous 
 integration tool Travis CI for more than 50 builds. All these features help to 
 build the notion that since we are building our analysis on top of a vast 
 dataset comprehending a wide diversity of projects, the threat of this analysis 
 not generalizing to other projects is mitigated.

\paragraph{Construct validity:}
 Travis job success is not the only way to analyze the success of a commit, and 
 that for different reasons the job may not be successful even if the code 
 commited does not have flaws (a flacky test case for example). Since continuous 
 integration is a best known practice in industry and the use of this methodology 
 spreads out into the open source community, we consider that cases of misuse of 
 CI such that it will lead to failing travis jobs when they should have actually 
 be successful are minimum, specially since having a solid build is one of the 
 main purposes of the methodology, and taking into account that the projects in 
 out dataset are projects with a history of using continuous integration and 
 having a certain level of popularity, this threat is minimized. We therefore 
 consider Travis job success to be a proxy 
 for success when integrating a commit into a larger project.

@@

\clg{per my email: fix the bibliography. Now.}
